Ornaments as indicators of social change before and after European contact at Kiwulan, Northeastern Taiwan

### 2019-08-20 analysis/paper/paper.Rmd
edits to abstract

---
title: "Ornaments as indicators of social change before and after European contact at Kiwulan, Northeastern Taiwan"
author:
  - Li-Ying Wang
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  The direct impact of European colonization since 16th century on local indigenous people has been widely discussed. In East Asia, direct European colonial rule on local communities was much less pronounced. However, the question of long-lasting indirect impacts on indigenous peoples in the periphery of colonial control remains largely unanswered. Using trade ornaments recovered from Kiwualn site (1400-1900 AD) in northeastern Taiwan, this research explores the indirect European colonial impacts on local societies by examining cultural context, variety, and spatial distribution of trade ornaments in terms of three time episodes relating different scales of culture contact. The results indicate the diversity of ornaments was greater and their spatial distribution was non-random during the European contact period that hints the increasing social inequality in a colonial context. This study demonstrates the use of foreign ornaments as a proxy to detect indirect colonial influence on local indigenous populations and inform on social changes relating to trade with colonial groups. Ornaments can give insights into the amplification of social inequality stimulated by European colonization. It also shows the agency of indigenous people to incorporate ornaments into their social system.
keywords: |
  Historical archaeology; Iron Age; European colonization; Trade ornaments; Exchange networks; Taiwan; Monte Carlo simulation
highlights: |
  1. Evidence of the quantity, variety, and spatial distribution of trade ornaments in both local cultural and colonial context is provided.
  2. Results of hypothesis testing point to greater amount and diversity of ornament types with non-random, and most highly concentrated spatial clusters during the European contact.
  3. Trade ornament as a prestige good may result in an increase in social inequality due to competition of wealth accumulation between individuals.
  4. Trade ornament can be a good proxy to detect indirect colonial influence on local indigenous populations and inform on social changes relating to trade with colonial groups.
---

>>>>

---
title: "Ornaments as indicators of social change before and after European contact at Kiwulan, Northeastern Taiwan"
author:
  - Li-Ying Wang
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  In East Asia, direct European colonial rule on local communities was much less pronounced that other parts of the world. Ij this region, the question of long-lasting indirect impacts on indigenous peoples in the periphery of colonial control remains largely unanswered. Using trade ornaments recovered from Kiwulan (1400-1900 AD) in northeastern Taiwan, I explore the indirect impacts of European colonial activties on local societies. I examine the cultural contexts, variety, and spatial distributions of trade ornaments in three time episodes relating different scales of culture contact. The results indicate the diversity of ornaments was greater and their spatial distribution was clustered during the European contact period. This hints at increasing social inequality in a colonial context. This study demonstrates the use of foreign ornaments as a proxy to detect indirect colonial influence on local indigenous populations and inform on social changes relating to trade with colonial groups. Ornaments can give insights into the amplification of social inequality stimulated by European colonization. It also shows the agency of indigenous people to incorporate ornaments into their social system.
keywords: |
  Historical archaeology; Iron Age; European colonization; Trade ornaments; Exchange networks; Taiwan; Monte Carlo simulation
highlights: |
  1. Evidence of the quantity, variety, and spatial distribution of trade ornaments in both local cultural and colonial context is provided.
  2. Results of hypothesis testing point to greater amount and diversity of ornament types with non-random, and most highly concentrated spatial clusters during the European contact.
  3. Trade ornament as a prestige good may result in an increase in social inequality due to competition of wealth accumulation between individuals.
  4. Trade ornament can be a good proxy to detect indirect colonial influence on local indigenous populations and inform on social changes relating to trade with colonial groups.
---
